---
title: "Prototype measurement framework: Triple diagnostic assessment in a single hospital visit for suspected breast cancer"
author: "Impact team"
date: "10 Oct 2022"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r TDA-rmd-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r TDA_setup, include = FALSE}

# Load relevant packages
library(dplyr)
library(lubridate)
library(plotly)

```

## Relevant guidance

The QS12 NICE quality standard on [Breast cancer](https://www.nice.org.uk/guidance/qs12) covers the care of people with breast cancer after they have been referred to a specialist team. The specific statement that we will measure here is:

> Quality statement 1 - People with suspected breast cancer referred to specialist services are offered the triple diagnostic assessment in a single hospital visit.

## Context

Early diagnosis of breast cancer allows for prompt treatment, which results in better health outcomes for people with breast cancer. Giving people with suspected breast cancer the triple diagnostic assessment at a single hospital visit will help to ensure rapid diagnosis. It will also help to reduce the anxiety and stress associated with multiple visits for different parts of the triple diagnostic assessment.

A triple diagnostic assessment comprises of the following three elements conducted in a single hospital visit:

- Clinical assessment
- Imaging - A mammogram and/or an ultrasound scan
- Biopsy - a tissue biopsy taken from the breast for histological examination by fine-needle aspiration or core biopsy.

The specific research question addressed in this prototype is as follows:

> What proportion of people diagnosed with breast cancer received a triple diagnostic assessment in a single hospital visit?

To answer this question, we will use data from the [National audit of breast cancer in older patients](https://www.nabcop.org.uk/reports/?filter_date=2022) (NABCOP) to assess uptake of triple diagnostic assessment. We will use this data to assess changes in uptake over time, as well as geographical variation in uptake. <!-- I would say "variation in uptake across trusts" rather than geographical variation, as you're not necessarily looking at region or city or ICB based variations. -->

## Methods

<!-- Maybe move data source first to make it clear that the actual analyses, defining of numerator and denominator, calculating of proportions etc was done be NABCOP, rather than us calculating it based on NABCOP published data? -->

### What is being measured? 
The proportion of breast cancer patients who received a triple diagnostic assessment within a single visit.

To determine whether a triple diagnostic assessment was received in a single visit, the following conditions have to be met:

- patient has a reported date of biopsy or cytology
- patient has a matching date of mammogram
- OR patient has no date for mammogram, but has a matching first seen date reported.

The numerator and denominator are as follows:

**Numerator**: Women in the denominator receiving triple diagnostic assessment in a single visit.  
**Denominator**: Women aged 50 and over diagnosed with non-screen detected early invasive breast cancer. 

<!-- I realise I haven't done this either but probably worth mentioning country: England (esp as NABCOP data includes Wales) -->

### Exclusions and caveats 
Women who were diagnosed with early invasive breast cancer during routine breast screening were excluded. This is because of established screening protocols where those with mammographic abnormalities are recalled for assessment and further imaging/biopsies, and thus would not have all tests done in the same visit. 

All men were excluded from analyses due to the low incidence making analyses unfeasible.

### Data source
[National Audit of Breast Cancer in Older Patients](https://www.nabcop.org.uk/reports/?filter_date=2022) (NABCOP). This audit uses data that has been provided by patients and collected by the NHS as part of their care and support. The data are collated, maintained and quality assured by the National Disease Registration Service (NDRS), which is part of NHS Digital (NHSD).

Data on triple diagnostic assessment is derived from the Cancer Outcomes and Services Dataset (COSD), a monthly data collection from service providers where items are submitted to the National Cancer Data Repository (NCDR) via multidisciplinary team electronic data collection systems.

Please refer to the [NABCOP Annual Report Methodology 2022](https://www.nabcop.org.uk/resources/nabcop-2022-annual-report-supplementary-materials/) for full details of how data was collected and processed. 

## Results

### National rate of triple diagnostic assessment

Among women aged 50 years and over diagnosed with non-screen detected early invasive breast cancer in 2019, 69% were calculated to have received triple diagnostic assessment (TDA) in a single visit. There was little variation from previous audit years, and between age groups.

```{r TDA-national}

df_nat <- readr::read_csv(here::here("data/NABCOP_audit_data_national.csv"),
                          col_types = "ccnn") %>%
          mutate(across(c(1, 2), ~lubridate::make_date(as.numeric(.x, 1, 1))))

# I think the code is wrong with as.numeric(.x, 1, 1) but it works because of the default arguments in make_date, where month = 1L and day = 1L

# as.numeric(.x, 1, 1) only returns the year as a numeric. See below:

  # as.numeric("2019", 1, 1)
  
  # test <- readr::read_csv(here::here("data/NABCOP_audit_data_national.csv"),
  #                           col_types = "ccnn") %>%
  #           mutate(across(c(1, 2), ~as.numeric(.x, 1, 1)))

# Can just do:
# df_nat <- readr::read_csv(here::here("data/NABCOP_audit_data_national.csv"),
#                           col_types = "ccnn") %>%
#           mutate(across(c(1, 2), ~lubridate::make_date(as.numeric(.x))))

# Also, I don't know if this is true as I don't use the here package but the docs (https://here.r-lib.org/articles/here.html) say that it's good practice to declare the location of the current script in the first chunk of the .Rmd with here::i_am(). Unsure if changes needed anyway as Rob removed the here calls as it doesn't function well when working outside the designated R project.

chart_title <- "<b>Percentage of women with non-screen detected early invasive breast cancer
                that received a triple diagnostic assessment in a single visit in England, 
                by year of diagnosis and age at diagnosis</b>"

df_nat %>% 
  plot_ly(x = ~diagnosis_year) %>% 
  add_lines(y = ~tda_perc_50_69,
            name = "50-69yrs",
            mode = "lines+markers",
            color = I("#228096"),
            marker = list(color = "#228096",
                          size = 8)) %>% 
  add_lines(y = ~tda_perc_70_plus,
            name = "70+yrs",
            color = I("#407291"),
            mode = "lines+markers",
            marker = list(color = "#00436C",
                          size = 8)) %>%
  layout(title = list(text = stringr::str_wrap(chart_title, width = 95),
                      font = list(size = 12),
                      x = 0.05,
                      y = 0.96,
                      yanchor = "top"),
         yaxis = list(title = "Percentage of women (%)",
                      ticksuffix = "%",
                      range = list(0,100)),
         xaxis = list(title = list(text = "Year of diagnosis",
                                   font = list(size=12),
                                   standoff = 10),
                      ticks = "outside",
                      nticks = 4),
         legend = list(xanchor = "center",
                       x = 0.82,
                       y = 0.95),
         hovermode = "x unified",
         margin = list(t = 65)) %>% 
  config(displaylogo = FALSE,
         modeBarButtonsToRemove = c("zoom", "pan", "select", "lasso"))
```

<!-- Nice graph. Is the reason why there are only data points from 2016, 2017 and 2019 vs 2014-2019) because of the availability of supplementary data from different audit years? May be worth clarifying as users may wonder why this data is less complete than Fig 3.4.1 in the NABCOP 2022 report. I'm guessing it's the same reason for collapsing 80+ with 70-79? Or small numbers. -->

### Variation in uptake by trust, 2019

<!-- 2019, not 2022? -->

There was marked variation by NHS organisation, with 37% of NHS organisations having less than 70% of patients estimated as receiving triple diagnostic assessment in a single visit.

```{r TDA-trust_chart}

df <- readr::read_csv(here::here("data/NABCOP_audit_data.csv"),
                      col_types = "ccccccnn",
                      na = c(""))

chart_title <- "<b>Percentage of women with non-screen detected early invasive breast cancer 
                (diagnosed in 2019)
                that received a triple diagnostic assessment in a single visit, 
                by diagnosing NHS organisation and age at diagnosis</b>"


df %>% 
  filter(diagnosis_year == "2019") %>% 
  arrange(desc(tda_perc_50_69)) %>% 
  plot_ly(x = ~reorder(ods_trust_nm, -tda_perc_50_69)) %>% 
  add_bars(y = ~tda_perc_50_69,
           name = "50-69yrs",
           color = I("#228096"),
           alpha = 0.6) %>%
  add_lines(y = ~tda_perc_70_plus,
            name = "70+yrs",
            color = I("#407291"),
            marker = list(color = "#00436C",
                          size = 4)) %>%
  layout(title = list(text = stringr::str_wrap(chart_title, width = 95),
                      font = list(size = 12),
                      x = 0.05,
                      yanchor = "top"),
         yaxis = list(title = "Percentage of women (%)",
                      ticksuffix = "%",
                      rangemode = "tozero"),
         xaxis = list(title = list(text = "NHS Organisation",
                                   font = list(size=12),
                                   standoff = 10),
                      showticklabels = FALSE,
                      ticks = "outside",
                      nticks = ~length(ods_trust_cd)),
         legend = list(xanchor = "center",
                       x = 0.85,
                       y = 0.9),
         hovermode = "x unified",
         margin = list(t = 50)) %>% 
  config(displaylogo = FALSE,
         modeBarButtonsToRemove = c("zoom", "pan", "select", "lasso"))
```

<!-- Agree with your point about adding n numbers if possible, though seems like it would require back calculations from the percentages provided in the supplementary materials. Not the best.

Seems like NABCOP have only done small number suppression on the denominator (Total women with early invasiove breast cancer not detected at screening in 2019 (N)), so would need to suppress if the back-calculated number who received TDA is between 1-9 (inclusive). -->

<!-- Maybe add a section on all the limitations around relying on NABCOP audit data to measure this rec, vs being able to do the analysis ourselves using NCRAS data? Issues around understanding the detailed methodology used by NABCOP (plus embedded assumptions), consistency in methodology across years, inability to look at subgroup analyses, confounders and data quality/completeness, and all the issues you highlighted in the Show and Tell slides. 

And maybe something on promising developments:
An item on whether triple diagnostic assessment happened in a single visit was added to COSD v9.0 and collected from September 2020 in England (pg 24, NABCOP 2022 report). Q: "Was a triple diagnostic assessment completed for the patient in a single visit, following initial referral? Yes/No/Not known"(pg 90, NABCOP 2022)

CancerStats shows relatively good completeness: "data reported on TDA in a single visit available for 47% of women aged 50+ years diagnosed between October 2020 and September 2021" (pg 72, NABCOP 2022) -->

## References

National Audit of Breast Cancer in Older Patients (NABCOP). NABCOP 2022 Annual Report. 2022. Available from: https://www.nabcop.org.uk/reports/nabcop-2022-annual-report/

<!-- Perhaps provide links in the references to the supplementary material Excel sheets which this data is from (e.g. Fig 2 https://www.nabcop.org.uk/content/uploads/2022/09/NABCOP_2022_Annual_Report_NHS_Organisation_Data_Viewer-Version15092022.xlsx), just for extra clarity. Not clear right now exactly where you got the data for the figures from. -->
