---
title: "Prototype measurement framework: Triple diagnostic assessment in a single hospital visit for suspected breast cancer"
author: "Impact team"
date: "10 Oct 2022"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r TDA-rmd-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

here::i_am("audit_prototype.Rmd")
```

```{r TDA_setup, include = FALSE}

# Load relevant packages
library(dplyr)
library(lubridate)
library(plotly)

```

## Relevant guidance

The QS12 NICE quality standard on [Breast cancer](https://www.nice.org.uk/guidance/qs12) covers the care of people with breast cancer after they have been referred to a specialist team. The specific statement that we will measure here is:

> Quality statement 1 - People with suspected breast cancer referred to specialist services are offered the triple diagnostic assessment in a single hospital visit.

## Context

Early diagnosis of breast cancer allows for prompt treatment, which results in better health outcomes for people with breast cancer. Giving people with suspected breast cancer the triple diagnostic assessment at a single hospital visit will help to ensure rapid diagnosis. It will also help to reduce the anxiety and stress associated with multiple visits for different parts of the triple diagnostic assessment.

A triple diagnostic assessment comprises of the following three elements conducted in a single hospital visit:

- Clinical assessment
- Imaging - A mammogram and/or an ultrasound scan
- Biopsy - a tissue biopsy taken from the breast for histological examination by fine-needle aspiration or core biopsy.

The specific research question addressed in this prototype is as follows:

> What proportion of people diagnosed with breast cancer received a triple diagnostic assessment in a single hospital visit?

To answer this question, we will use data from the [National audit of breast cancer in older patients](https://www.nabcop.org.uk/reports/?filter_date=2022) (NABCOP) to assess uptake of triple diagnostic assessment. We will use this data to assess changes in uptake over time, and variation in uptake across trusts.

## Methods

### Data source
[National Audit of Breast Cancer in Older Patients](https://www.nabcop.org.uk/reports/?filter_date=2022) (NABCOP). This audit uses data that has been provided by patients and collected by the NHS as part of their care and support. The data are collated, maintained and quality assured by the National Disease Registration Service (NDRS), which is part of NHS Digital (NHSD).

Data on triple diagnostic assessment is derived from the Cancer Outcomes and Services Dataset (COSD), a monthly data collection from service providers where items are submitted to the National Cancer Data Repository (NCDR) via multidisciplinary team electronic data collection systems.

Please refer to the [NABCOP Annual Report Methodology 2022](https://www.nabcop.org.uk/resources/nabcop-2022-annual-report-supplementary-materials/) for full details of how data was collected and processed. 


### What is being measured? 
The proportion of breast cancer patients who received a triple diagnostic assessment within a single visit.

To determine whether a triple diagnostic assessment was received in a single visit, the following conditions have to be met:

- patient has a reported date of biopsy or cytology
- patient has a matching date of mammogram
- OR patient has no date for mammogram, but has a matching first seen date reported.

The numerator and denominator are as follows:

**Numerator**: Women in the denominator receiving triple diagnostic assessment in a single visit.  
**Denominator**: Women aged 50 and over diagnosed with non-screen detected early invasive breast cancer in England. 

### Exclusions and caveats 
Women who were diagnosed with early invasive breast cancer during routine breast screening were excluded. This is because of established screening protocols where those with mammographic abnormalities are recalled for assessment and further imaging/biopsies, and thus would not have all tests done in the same visit. 

All men were excluded from analyses due to the low incidence making analyses unfeasible.


## Results

### National rate of triple diagnostic assessment

Among women aged 50 years and over diagnosed with non-screen detected early invasive breast cancer in 2019, 69% were calculated to have received triple diagnostic assessment (TDA) in a single visit. There was little variation from previous audit years, and between age groups. Audit years prior to 2019 were not included due to changes in the audit specification for triple diagnostic assessment, making earlier years difficult to compare. While the 2022 audit includes charts showing uptake by diagnosis year dating back to 2014, only the most recent year of data was published.

```{r TDA-national}

df_nat <- readr::read_csv(here::here("data/NABCOP_audit_data_national.csv"),
                          col_types = "ccnn") %>%
          mutate(across(c(1, 2), ~lubridate::make_date(as.numeric(.x), 1L, 1L)))

chart_title <- "<b>Percentage of women with non-screen detected early invasive breast cancer
                that received a triple diagnostic assessment in a single visit in England, 
                by year of diagnosis and age at diagnosis</b>"

df_nat %>% 
  plot_ly(x = ~diagnosis_year) %>% 
  add_lines(y = ~tda_perc_50_69,
            name = "50-69yrs",
            mode = "lines+markers",
            color = I("#228096"),
            marker = list(color = "#228096",
                          size = 8)) %>% 
  add_lines(y = ~tda_perc_70_plus,
            name = "70+yrs",
            color = I("#407291"),
            mode = "lines+markers",
            marker = list(color = "#00436C",
                          size = 8)) %>%
  layout(title = list(text = stringr::str_wrap(chart_title, width = 95),
                      font = list(size = 12),
                      x = 0.05,
                      y = 0.96,
                      yanchor = "top"),
         yaxis = list(title = "Percentage of women (%)",
                      ticksuffix = "%",
                      range = list(0,100)),
         xaxis = list(title = list(text = "Year of diagnosis",
                                   font = list(size=12),
                                   standoff = 10),
                      ticks = "outside",
                      nticks = 4),
         legend = list(xanchor = "center",
                       x = 0.82,
                       y = 0.95),
         hovermode = "x unified",
         margin = list(t = 65)) %>% 
  config(displaylogo = FALSE,
         modeBarButtonsToRemove = c("zoom", "pan", "select", "lasso"))
```

### Variation in uptake by trust, 2019

When examining the most recent year of data, including only women who received a diagnosis in 2019, we observed marked variation in uptake across NHS organisations, with 37% having less than 70% of patients estimated as receiving triple diagnostic assessment in a single visit.

```{r TDA-trust_chart}

df <- readr::read_csv(here::here("data/NABCOP_audit_data.csv"),
                      col_types = "ccccccnn",
                      na = c(""))

chart_title <- "<b>Percentage of women with non-screen detected early invasive breast cancer 
                (diagnosed in 2019)
                that received a triple diagnostic assessment in a single visit, 
                by diagnosing NHS organisation and age at diagnosis</b>"


df %>% 
  filter(diagnosis_year == "2019") %>% 
  arrange(desc(tda_perc_50_69)) %>% 
  plot_ly(x = ~reorder(ods_trust_nm, -tda_perc_50_69)) %>% 
  add_bars(y = ~tda_perc_50_69,
           name = "50-69yrs",
           color = I("#228096"),
           alpha = 0.6) %>%
  add_lines(y = ~tda_perc_70_plus,
            name = "70+yrs",
            color = I("#407291"),
            marker = list(color = "#00436C",
                          size = 4)) %>%
  layout(title = list(text = stringr::str_wrap(chart_title, width = 95),
                      font = list(size = 12),
                      x = 0.05,
                      yanchor = "top"),
         yaxis = list(title = "Percentage of women (%)",
                      ticksuffix = "%",
                      rangemode = "tozero"),
         xaxis = list(title = list(text = "NHS Organisation",
                                   font = list(size=12),
                                   standoff = 10),
                      showticklabels = FALSE,
                      ticks = "outside",
                      nticks = ~length(ods_trust_cd)),
         legend = list(xanchor = "center",
                       x = 0.85,
                       y = 0.9),
         hovermode = "x unified",
         margin = list(t = 50)) %>% 
  config(displaylogo = FALSE,
         modeBarButtonsToRemove = c("zoom", "pan", "select", "lasso"))
```

<!-- Maybe add a section on all the limitations around relying on NABCOP audit data to measure this rec, vs being able to do the analysis ourselves using NCRAS data? Issues around understanding the detailed methodology used by NABCOP (plus embedded assumptions), consistency in methodology across years, inability to look at subgroup analyses, confounders and data quality/completeness, and all the issues you highlighted in the Show and Tell slides. 

And maybe something on promising developments:
An item on whether triple diagnostic assessment happened in a single visit was added to COSD v9.0 and collected from September 2020 in England (pg 24, NABCOP 2022 report). Q: "Was a triple diagnostic assessment completed for the patient in a single visit, following initial referral? Yes/No/Not known"(pg 90, NABCOP 2022)

CancerStats shows relatively good completeness: "data reported on TDA in a single visit available for 47% of women aged 50+ years diagnosed between October 2020 and September 2021" (pg 72, NABCOP 2022) -->

## References

National Audit of Breast Cancer in Older Patients (NABCOP). NABCOP 2022 Annual Report. 2022. Available from: https://www.nabcop.org.uk/reports/nabcop-2022-annual-report/

National Audit of Breast Cancer in Older Patients (NABCOP) supplementary materials . NABCOP Annual Report Methodology 2022. Available from:https://www.nabcop.org.uk/resources/nabcop-2022-annual-report-supplementary-materials/

National Audit of Breast Cancer in Older Patients (NABCOP) supplementary materials. NABCOP Annual Report 2022 NHS Organisation Data Viewer. Available from:https://www.nabcop.org.uk/resources/nabcop-2022-annual-report-supplementary-materials/
